---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# ivx: Robust Econometric Inference

[![CRAN status](https://www.r-pkg.org/badges/version/ivx)](https://cran.r-project.org/package=ivx)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/kvasilopoulos/ivx.svg?branch=master)](https://travis-ci.org/kvasilopoulos/ivx)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/kvasilopoulos/ivx?branch=master&svg=true)](https://ci.appveyor.com/project/kvasilopoulos/ivx)


Conducting inference on the regression coefficient with a highly persistent regressor, using the the IVX approach for univariate (Magdalinos and Phillips, 2009b) and multivariate (Kostakis Magdalinos and Stamatogiannis, 2015) regression. Testing can be used for long-horizon  predictability as well.
    

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("kvasilopoulos/ivx")
```
## Usage

Load ivx and dplyr for data managment

```{r message = FALSE}
library(ivx)
```

This is a basic example, lets load the data first:

```{r example}
# Monthly data from Kostakis et al (2014)
monthly %>%
  names()

```

## Univariate

And then do the univariate estimation:

```{r uni}
ivx(Ret ~ DP, data = monthly) %>% 
  summary()

ivx(Ret ~ DP, data = monthly, horizon = 4) %>% 
  summary()

```

## Multivariate

And the multivariate estimation, for one or multiple horizons:

```{r multi}
ivx(Ret ~ DP + TBL, data = monthly) %>% 
  summary()

ivx(Ret ~ DP + TBL, data = monthly, horizon = 4) %>% 
  summary()

```

---

Please note that the 'ivx' project is released with a [Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
