---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# ivx: Predictive Regressions

 [![Travis build status](https://travis-ci.org/kvasilopoulos/ivx.svg?branch=master)](https://travis-ci.org/kvasilopoulos/ivx)

The goal of ivx is to offer robust econometric inference for predictive regressions.

![first equation](https://latex.codecogs.com/gif.latex?%5Cbegin%7Balign*%7D%20y_t%20%26%3D%20%5Cbeta%20y_%7Bt-1%7D%20&plus;%20%5Cepsilon_t%5C%5C%20x_t%20%26%3D%20%5Crho%20x_%7Bt-1%7D%20&plus;%20u_t%20%5Cend%7Balign*%7D)


### Solution methods

* The Bonferroni method
  - Cavanagh et al (1995)
  - Campbell and Yogo (2006)
* A conditional likelihood approach
  - Jansson and Moreira (2006)
* A control function approach
  - Elliot (2001)
* The IVX approach
  - Magdalinos and Phillips (2009b)

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("kvasilopoulos/ivx")
```
## Example

Load ivx and dplyr for data managment
```{r message = FALSE}
library(ivx)
library(dplyr)
```

This is a basic example, lets load the data first:
```{r example}
# Monthly data from Kostakis et al (2014)
datam %>%
  names()

univariate <- datam %>%
  select(Ret, D_P)

multivariate <- datam %>%
  select(Ret, D_P, TBL)

```

And then do the univariate estimation:
```{r uni}
ivx(Ret ~ D_P, data = univariate) %>% 
  summary()

qtest(univariate$Ret, univariate$D_P)
```

And the multivariate estimation, for one or multiple horizons:
```{r multi, eval=FALSE, include=FALSE}
ivx(Ret ~ D_P + TBL, data = multivariate) %>% 
  summary()

ivx(Ret ~ D_P + TBL, data = multivariate, horizon = 4) %>% 
  summary()

```
