---
title: "Non-IVX benchmarks: Bonferroni Q, ARM, hybrid t, empirical likelihood"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Non-IVX benchmarks: Bonferroni Q, ARM, hybrid t, empirical likelihood}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup, message = FALSE}
library(ivx)
```

Five non-IVX procedures are included as the benchmarks the IVX literature
compares against: the Bonferroni Q-test of Campbell & Yogo (2006), `cy_test()`;
the augmented regression method of Amihud, Hurvich & Wang (2009), `arm()`;
the control-function regression of Elliott (2011), `elliott_cf()`; the hybrid
switching t-test of Harvey, Leybourne & Taylor (2021), `hlt_test()`; and the
unified empirical likelihood test of Liu, Yang, Cai & Peng (2019), `el_test()`.

# Bonferroni Q-test

`cy_test()` is the Campbell & Yogo (2006) procedure, the feasible version of
the sup-bound / Bonferroni idea of Cavanagh, Elliott & Stock (1995). If
$\rho$ were known, the UMP conditional test of $\beta = 0$ is the t-ratio of
the regression augmented with $x_t - \rho x_{t-1}$ (the "Q-test"), whose
estimate
$$
\hat\beta(\rho) = \frac{\sum_t x^\mu_{t-1}\big(y_t - \tfrac{\sigma_{ue}}{\sigma_e\omega}(x_t - \rho x_{t-1})\big) - \tfrac{T}{2}\tfrac{\sigma_{ue}}{\sigma_e\omega}(\omega^2 - \sigma_v^2)}{\sum_t x^{\mu 2}_{t-1}}
$$
(their eq. 25, with the AR($p$) correction of Appendix A) has standard error
$\sigma_u\sqrt{1-\delta^2}/(\sum x^{\mu 2}_{t-1})^{1/2}$. Since $\rho$ is
not consistently estimable, a confidence interval for it is obtained by
inverting the DF-GLS statistic (Stock, 1991; Elliott, Rothenberg & Stock,
1996) and the Bonferroni interval for $\beta$ runs from
$\hat\beta(\bar\rho) - 1.645\,se$ to $\hat\beta(\underline\rho) + 1.645\,se$.
The levels of the interval for $\rho$ come from the paper's Table 2, which
tightens the plain Bonferroni bound so that the one-sided test has size
exactly 5% for some $c$.

The DF-GLS null quantiles as a function of $c$ are simulated once
(`data-raw/dfgls-quantiles.R`, $c \in [-100, 10]$, 20 000 replications of a
600-step OU process) and shipped as internal data; the 5% quantile at
$c = 0$ reproduces the $-1.95$ of Elliott et al. (1996).

```{r}
cy_test(Ret ~ DP, data = kms)
cy_test(Ret ~ EP, data = kms)
```

In a Monte Carlo with $T = 200$ and 500 replications, the right-tailed 5%
Q-test rejects a true null 4.4–5.8% of the time and the left-tailed one
1.2–3.6% for $\delta \in \{-0.9, -0.5\}$ and $c \in \{0, -20\}$ — the
asymmetry the paper describes (Section 3.4: left-tailed probability "can be
as small as 1.2%").


# Augmented regression method

`arm()` is reduced-bias OLS for *stationary* persistent predictors
(Demetrescu & Rodrigues, 2022, build `ivx_ra()` on the same augmentation idea).

## Idea

With $y_t = \alpha + \beta' x_{t-1} + u_t$ and $x_t = \mu + \Phi x_{t-1} + v_t$,
write $u_t = \phi' v_t + e_t$ with $e_t$ independent of $x$. OLS of $y_t$ on
$x_{t-1}$ is biased because $u_t$ is correlated with $v_t$ and hence with the
future of the regressor (Stambaugh, 1999). Regressing $y_t$ on $x_{t-1}$ *and*
$v_t$ removes it — if $v_t$ were known. Using OLS VAR residuals does not help
(they are orthogonal to $x_{t-1}$ by construction), so the VAR coefficient
matrix is first bias-corrected with the Nicholls & Pope (1988) expansion
$$
E[\hat\Phi - \Phi] = -\tfrac{1}{n}\,\Sigma_v\Big[(I-\Phi')^{-1} +
\Phi'(I-\Phi'^2)^{-1} + \sum_{\lambda \in \mathrm{Spec}(\Phi)} \lambda (I-\lambda\Phi')^{-1}\Big]\Gamma_0^{-1},
$$
iterated up to `iter` times (Kendall's $1 + 3\phi$ in the scalar case), and
the corrected residuals $v^c_t = x_t - \hat\mu^c - \hat\Phi^c x_{t-1}$ are the
control variables.

The OLS covariance of the augmented regression is wrong for $\hat\beta^c$
because $v^c_t$ is estimated. The paper's estimator (eqs 7–8) adds the
uncertainty of $\hat\Phi^c$:
$$
\widehat{\mathrm{cov}}[\hat\beta^c] = (\hat\phi'\hat\Sigma_v\hat\phi)\,(X'X)^{-1}_{xx}
+ \hat\sigma_e^2 \Big[\tfrac{\sum_t r_{it} r_{jt}}{\sum_t r_{it}^2 \sum_t r_{jt}^2}\Big]_{ij},
$$
with $r_{jt}$ the residual of $x_{j,t-1}$ on the other regressors of the
augmented regression. `summary()` reports $t$-ratios and the joint Wald test
from this matrix.

```{r}
m <- arm(Ret ~ DP + TBL, data = kms)
summary(m)
m$Phi      # bias-corrected VAR(1) coefficients
m$Phi_ols  # OLS
```

## Check

With $\phi = 0.9$, $\mathrm{corr}(u, v) = -0.9$, $n = 100$ and $\beta = 0$
the OLS slope has mean bias about $0.037$ ($\approx -\rho(1+3\phi)/n$) over
400 replications; `arm()` reduces it to $0.003$, with $\hat\Phi^c$ averaging
$0.894$ against an OLS mean of $0.857$. In the paper's two-predictor Case 1 the
5% $t$-tests reject 7–8% at $n = 100$, in line with the 6–10% ($n = 50$) and
5–7% ($n = 200$) the paper reports.

## Caveats

* Assumes stationary predictors: the correction diverges as the largest
  eigenvalue of $\Phi$ approaches one, and the iteration stops if the
  corrected VAR becomes non-stationary. For near-unit-root predictors use
  `ivx()` or `ivx_ra()`.
* VAR(1) only, short horizon only, no bootstrap.

# Control-function regression

Elliott (2011) shows that if stationary covariates $z_t$ are available that
are contemporaneously correlated with the shocks to both the predictor and
the response, the Wald test of $eta = 0$ in
$$
y_t = lpha + eta' x_{t-1} + \gamma' Z_t + 	ilde u_t, \qquad
Z_t = (z_t', z_{t-1}', \dots, z_{t-q}')',
$$
is asymptotically $\chi^2$ whatever the persistence of $x_t$, because the
covariates "orthogonalise" the innovations (his Theorem 2 versus the
Elliott–Stock 1994 distribution of Theorem 1). The covariates are a modelling
choice, not a data construction — `arm()` and `ivx_ra()` are the feasible
versions where the control variable is built from the predictor's own
innovations. `elliott_cf()` runs the regression, reports the Wald test with
Eicker-White standard errors and the correlation that remains between the
regression residuals and the predictor innovations, which should be near zero.

```{r}
elliott_cf(Ret ~ DP, ~ TBL, data = kms, lags = 1)
```

With a covariate that does absorb the correlation ($z_t = v_t + $ noise,
$\delta = -0.9$, $c = 0$, $T = 200$) the 5% test rejects a true null 4% of the
time against 23% for plain OLS (500 replications).

# Hybrid t-test

`hlt_test()` implements $T_{hyb}$ of Harvey, Leybourne & Taylor (2021) for a
single predictor. It uses only regression t-ratios: the standard one, $T$,
and $	ilde T$, in which the predictor is quasi-GLS demeaned with
$ar c = 7$ as in Elliott et al. (1996). Under strong persistence their null
distributions depend on $c$ and on $
ho_{xy}$, so the paper tabulates
*conservative* critical values — the maximum over $c$ of the $lpha$-level
quantile — as response surfaces in $\hat
ho_{xy}$ (their Table 1), and
switches:

1. if the ADF normalised-bias statistic (lag length by MBIC) is below
   $-4\sqrt{T}$ the predictor is weakly persistent: $T$ with a normal
   critical value;
2. otherwise, for an upper-tail test, $T$ with $cv(\hat
ho_{xy})$ when
   $\hat
ho_{xy} > -0.1$ and $	ilde T$ with $	ilde{cv}(\hat
ho_{xy})$
   when $\hat
ho_{xy} < -0.1$ (quasi-GLS demeaning pays off only when the
   endogeneity is strong); lower-tail tests mirror this.

```{r}
hlt_test(Ret ~ DP, data = kms)
hlt_test(Ret ~ TBL, data = kms, alternative = "less", level = 0.1)
```

In the paper's design ($T = 200$, upper-tail 5% test) the empirical size in
1000 replications is 0.07/0.04/0.05 for $
ho_{xy} = -0.9$ and
$
ho = 1, 0.95, 0.5$, and 0.00/0.02/0.04 for $
ho_{xy} = 0.9$ — the
"undersize for positive $
ho_{xy}$ in the more persistent cases" the paper
notes is the price of the conservative critical values. The test is one-sided
by construction and has no p-value; the object reports the selected test, its
statistic and critical value.

# Unified empirical likelihood test

`el_test()` implements Liu, Yang, Cai & Peng (2019). Their model adds the
lagged difference of the predictor,
$$
Y_t = lpha + eta_1 \Delta X_{t-1} + eta_2 X_{t-2} + U_t,
$$
so that $Y_t$ can be stationary whether or not $X_t$ is. The intercept is
removed by differencing at lag $m = \lfloor n/2 
floor$ (Zhu, Cai & Peng,
2014), and the empirical likelihood is built on the two score equations, the
second weighted by $1/\sqrt{1 + 	ilde X_{t-2}^2}$ so that its sample
variance converges whatever the persistence. The profile EL ratios for
$eta_2 = 0$ (no predictability), $eta_1 = 0$ and the joint null are then
$\chi^2(1)$, $\chi^2(1)$, $\chi^2(2)$ with no tuning parameter and no
persistence classification (Theorem 2). The price is efficiency: only $m - 2$
differenced observations enter, and the paper reports that the test on
$eta_1$ is oversized at $n = 200$.

```{r}
el_test(Ret ~ DP, data = kms)
```

The EL dual problem is solved by Newton's method with Owen's (2001)
pseudo-logarithm; the profiles are minimised over the nuisance coefficient
by BFGS from the OLS start.

## References

* Amihud, Y., Hurvich, C. M., & Wang, Y. (2009). Multiple-predictor
  regressions: Hypothesis testing. *Review of Financial Studies*, 22(1),
  413–434.
* Amihud, Y., & Hurvich, C. M. (2004). Predictive regressions: A reduced-bias
  estimation method. *Journal of Financial and Quantitative Analysis*, 39(4),
  813–841.
* Campbell, J. Y., & Yogo, M. (2006). Efficient tests of stock return
  predictability. *Journal of Financial Economics*, 81(1), 27–60.
* Cavanagh, C. L., Elliott, G., & Stock, J. H. (1995). Inference in models
  with nearly integrated regressors. *Econometric Theory*, 11(5), 1131–1147.
* Elliott, G. (2011). A control function approach for testing the usefulness
  of trending variables in predictive regressions and econometric models.
  *Journal of Econometrics*, 164(1), 79–91.
* Elliott, G., Rothenberg, T. J., & Stock, J. H. (1996). Efficient tests for
  an autoregressive unit root. *Econometrica*, 64(4), 813–836.
* Harvey, D. I., Leybourne, S. J., & Taylor, A. M. R. (2021). Simple tests
  for stock return predictability with good size and power properties.
  *Journal of Econometrics*, 224(1), 198–214.
* Liu, X., Yang, B., Cai, Z., & Peng, L. (2019). A unified test for
  predictability of asset returns regardless of properties of predicting
  variables. *Journal of Econometrics*, 208(1), 141–159.
* Nicholls, D. F., & Pope, A. L. (1988). Bias in the estimation of
  multivariate autoregressions. *Australian Journal of Statistics*, 30A,
  296–309.
* Owen, A. B. (2001). *Empirical Likelihood*. Chapman & Hall.
* Stock, J. H. (1991). Confidence intervals for the largest autoregressive
  root in U.S. macroeconomic time series. *Journal of Monetary Economics*,
  28(3), 435–459.
* Stambaugh, R. F. (1999). Predictive regressions. *Journal of Financial
  Economics*, 54(3), 375–421.
