---
title: "IV tests with combined instruments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IV tests with combined instruments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup, message = FALSE}
library(ivx)
```

`ivx_iv()` implements the instrumental-variable predictability tests of
Breitung & Demetrescu (2015). IVX is one member of the family: any instrument
$z_{t-1}$ that is less persistent than $x_{t-1}$ but correlated with it gives a
2SLS $t$-ratio with a standard normal null limit whatever the local-to-unity
parameter $c$.

## Instruments

Type-I instruments are functions of the predictor:

* `"frac"`: the truncated fractional difference $\Delta_+^{d} x_{t-1}$,
  $d = 1/2$ by default — $I(1/2)$ when $x$ is $I(1)$, still informative when
  $x$ is stationary;
* `"diff"`: the long difference $x_{t-1} - x_{t-1-k_T}$ with
  $k_T = \lfloor 0.2\,T^{0.85} \rfloor$;
* the IVX filter of Kostakis et al. (2015) is the paper's "mild integration"
  case (`ivx()`).

Type-II instruments are deterministic and only correlate with a
near-integrated predictor: `"sin"` uses $\sin(\pi t/T)$ (with $K$
predictors, frequencies $k = 1, \dots, K$). They are weak when $x$ is
stationary, so the paper recommends the over-identified combination
`"comb"` = sine + fractional: the 2SLS statistic is asymptotically driven by
whichever instrument is informative (Theorem 3), so it keeps power in both
regimes.

## The statistic

With $Z$ the instrument matrix (intercept included in both stages) and
$\hat u_t$ the OLS residuals of the predictive regression, the squared
$t$-ratio of eq. (12) is
$$
t^2 = \frac{\big(x'Z(Z'Z)^{-1}Z'y\big)^2}
{x'Z(Z'Z)^{-1}\big(\sum_t z_t z_t' \hat u_t^2\big)(Z'Z)^{-1}Z'x} \;\to\; \chi^2(1),
$$
and its $K$-predictor Wald analogue is $\chi^2(K)$ (Theorem 7).

```{r}
summary(ivx_iv(Ret ~ DP + TBL, data = kms))
sapply(c("comb", "sin", "frac", "diff"),
       function(i) ivx_iv(Ret ~ DP, data = kms, instruments = i)$tstat)
```

## Replication of Table 1

The paper's design: $T = 250$, $\mathrm{corr}(u, v) = 0.9$,
$\beta = b/T$, two-sided tests at 10%. Rejection rates in 1000
replications (paper values in brackets, 10 000 replications):

| $\rho$ | $b$ | comb | sin | frac | diff |
|---|---|---|---|---|---|
| 1.00 | 0 | 12.7 (11.2) | 11.5 (9.9) | 12.2 (11.1) | 11.2 (12.5) |
| 1.00 | 10 | 52.6 (65.7) | 38.5 (61.4) | 41.5 (33.4) | 30.4 (33.6) |
| 1.00 | 20 | 88.5 (91.3) | 59.6 (79.5) | 81.0 (66.9) | 68.9 (66.1) |
| 0.96 | 0 | 11.6 (10.4) | 9.7 (9.9) | 11.4 (9.4) | 10.7 (10.5) |
| 0.96 | 10 | 42.6 (47.3) | 29.4 (34.4) | 31.4 (30.1) | 33.2 (33.4) |

Sizes and the ranking (combination best, sine strong only near the unit
root) reproduce; the split of power between the sine and fractional
instruments differs from the paper, which does not report the exact
demeaning and initialisation used for the deterministic instrument.

## Caveats

* Short horizon only, no bootstrap.
* The sign (Cauchy) instrument of the paper needs recursive demeaning of
  the regressor and forward demeaning of the response and is not included.
* Several type-II instruments per predictor are not allowed (Assumption 5);
  the implementation uses one sine frequency per predictor.

## References

* Breitung, J., & Demetrescu, M. (2015). Instrumental variable and variable
  addition based inference in predictive regressions. *Journal of
  Econometrics*, 187(1), 358–375.
