---
title: "Residual-augmented IVX"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Residual-augmented IVX}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup, message = FALSE}
library(ivx)
```

`ivx_ra()` implements the residual-augmented IVX estimator of Demetrescu &
Rodrigues (2022), a bias-reduced version of IVX in the spirit of Amihud &
Hurvich (2004).

## Idea

Write the innovations of the predictor's autoregression as $v_t$ and
decompose the regression error as $u_t = \gamma' v_t + \eta_t$ with
$\eta_t$ uncorrelated with $v_t$. The finite-sample bias of any estimator of
$\beta$ comes from $\gamma \ne 0$. If $v_t$ were observed, adding it as a
regressor,

$$
\tilde y_t = \beta' \tilde x_{t-1} + \gamma' v_t + \eta_t,
$$

would remove the problem. Amihud & Hurvich estimate $v_t$ from a
bias-corrected autoregression; Demetrescu & Rodrigues show that augmenting the
*IVX* regression with plain autoregressive residuals is enough to reduce the
bias substantially while retaining the persistence-robust $\chi^2$ inference.

## Procedure

1. Fit an autoregression of order $p$ in levels to the predictors (a VAR
   with several predictors), **without intercept**, $p$ chosen by AIC (the
   paper's recommendation); keep the residuals $\hat\varepsilon_t$,
   $t = p+1, \dots, n$, and demean them.
2. Regress the demeaned response on $\hat\varepsilon_t$ by OLS and keep the
   residuals $\tilde y_t = \tilde y_t^{\,0} - \hat\gamma' \hat\varepsilon_t$.
3. Estimate $\beta$ by IVX of $\tilde y_t$ on $\tilde x_{t-1}$ with the KMS
   instrument (not demeaned).

Standard errors are the heteroskedasticity-robust form of the paper's
eq. (9)/(14), which stay valid whether $x_t$ is stationary or near-integrated:

$$
\widehat{\mathrm{Cov}}(\tilde\beta_{ivx}) = B^{-1} M B^{-1\prime}, \quad
B = \sum_t z_{t-1}\tilde x_{t-1}', \quad
M = \sum_t z_{t-1} z_{t-1}' \tilde\varepsilon_t^2 + \hat Q_T - n \bar z \bar z' \hat\Sigma_{FM},
$$

where $\tilde\varepsilon_t$ are the OLS residuals of the augmented
regression, $\hat Q_T = H_{zx} H_{xx}^{-1} \tilde H_{xx} H_{xx}^{-1} H_{zx}'$
accounts for the estimation of $\hat\varepsilon_t$ (it matters only in the
stationary case), and the last term is the Kostakis et al. (2015) intercept
correction that the paper also uses in its simulations.

```{r}
m <- ivx_ra(Ret ~ DP + TBL, data = kms)
m
summary(m)
```

Compared with plain IVX the point estimate on the strongly endogenous
predictor (`DP`, $\delta \approx -0.98$) shrinks and the standard error is
smaller — the efficiency gain from removing the $\gamma' v_t$ component of the
error.

```{r}
rbind(ivx = coef(ivx(Ret ~ DP + TBL, data = kms)), ivx_ra = coef(m))
```

## Replication: size in the paper's Monte Carlo

The paper has no reproducible empirical application, so the check is against
the size rows of its Table 3 (right-sided 5% tests, $T = 200$, DGP (24)–(25)
with $\rho = 1 - c/T$, short-run AR parameter $-0.5$ and innovation
correlation $-0.95$). With 10 000 replications:

| $c$ | paper $t_{ivx}$ | `ivx` | paper $\tilde t_{ivx}^{\mu_0}$ | `ivx_ra` |
|---|---|---|---|---|
| 0  | 0.116 | 0.129 | 0.054 | 0.064 |
| 10 | 0.088 | 0.089 | 0.055 | 0.064 |
| 20 | 0.074 | 0.072 | 0.055 | 0.064 |
| 30 | 0.066 | 0.057 | 0.053 | 0.055 |
| 40 | 0.064 | 0.057 | 0.050 | 0.059 |
| 50 | 0.061 | 0.051 | 0.050 | 0.056 |

The over-rejection of plain IVX under strong endogeneity is largely removed;
the residual offset of about 0.01 at small $c$ is shared by the plain IVX
column and so reflects a detail of the simulation design rather than the
estimator.

## Caveats and implementation notes

* Two details the paper leaves implicit are decisive: the autoregression must
  be fitted **without** an intercept (the paper's $\mu_0$ variant; its
  OLS-demeaned $\mu_1$ variant is much more biased and "not recommended for
  testing"), and the residuals must be **demeaned** before augmentation —
  without that the term $\gamma \bar{\hat\varepsilon}$ survives and the bias
  reduction fails. Both are built in and neither is optional.
* The remaining bias is $-\gamma$ times the bias of the AR coefficients, so
  it grows with $|\gamma|$ and with the persistence of the predictor; it is
  much smaller than for plain IVX but not zero.
* Short horizon only, and no bootstrap: `ivx_boot()` does not accept
  `ivx_ra` objects.
* The AR order is selected by AIC in levels, which the authors argue copes
  with both the stationary and the integrated case; `ar` fixes it instead.

## References

* Demetrescu, M., & Rodrigues, P. M. M. (2022). Residual-augmented IVX
  predictive regression. *Journal of Econometrics*, 227(2), 429–460.
* Amihud, Y., & Hurvich, C. M. (2004). Predictive regressions: A reduced-bias
  estimation method. *Journal of Financial and Quantitative Analysis*, 39(4),
  813–841.
