---
title: "Robust inference: Eicker-White and wild bootstrap"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Robust inference: Eicker-White and wild bootstrap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup, message = FALSE}
library(ivx)
```

Demetrescu, Georgiev, Rodrigues & Taylor (2023, *DGRT*) extend the IVX tests
of Kostakis et al. (2015) in two directions implemented here: validity under
general heteroskedasticity through Eicker–White standard errors
(`ivx(robust = TRUE)`, see `vignette("ivx")`), and bootstrap implementations
that markedly improve finite-sample size, in particular for one-sided tests
and with several predictors (`ivx_boot()`).

## Why bootstrap

Even with a valid $\chi^2$ limit, the IVX t-ratio is biased in finite
samples when the predictor is strongly persistent and $\delta$ is large in
absolute value: the estimate of the intercept interacts with the instrument,
and the leading higher-order term shifts the statistic (Hosseinkouchack &
Demetrescu 2021). With $T = 250$, $\delta = -0.95$ and a unit-root predictor,
DGRT report empirical sizes of 10–17% for the right-tailed 5% test.

## The two schemes

Let $\hat u_t$ be the OLS residuals of the predictive regression and
$R_t$ i.i.d. multipliers with mean zero and unit variance (Rademacher by
default, `dist = "normal"` optional).

**Fixed regressor wild bootstrap (FRWB)**, DGRT Algorithm 2. The bootstrap
response imposes the null, $y_t^* = R_t \hat u_t$; the predictors and the
instrument are kept fixed. The IVX statistics are recomputed on
$(y_t^*, x_{t-1}, z_{t-1})$.

**Residual wild bootstrap (RWB)**, DGRT Algorithm 1 and Remark 22. Fit an
autoregression in levels to the predictors (a VAR with several predictors),

$$
x_t = \hat m + \sum_{j=1}^{p+1} \hat a_j x_{t-j} + \hat v_t,
$$

with the order chosen by BIC (Remark 24). Draw
$(u_t^*, v_t^*) = (R_t \hat u_t, R_t \hat v_t)$ with the *same* $R_t$ so the
endogeneity correlation is preserved, rebuild
$x_t^* = \sum_j \hat a_j x_{t-j}^* + v_t^*$ from zero initial conditions and
$z_t^*$ from $\Delta x_t^*$ with the original $\rho_n$, and recompute the
statistics on $(y_t^*, x_{t-1}^*, z_{t-1}^*)$. RWB is the scheme that controls
size under strong persistence; FRWB is simpler and needs no lag choice.

The p-value is $1 - G^*(\text{stat})$, the bootstrap tail probability, for the
joint and individual Wald statistics and for the t-ratios against
$\beta < 0$, $\beta > 0$ and $\beta \ne 0$.

```{r}
mod <- ivx(Ret ~ DP + TBL, data = kms)
ivx_boot(mod, B = 499, type = "rwb", seed = 1)
```

The asymptotic joint p-value is `r round(1 - pchisq(mod$Wald_Joint, 2), 3)`;
the bootstrap moves it up, the direction DGRT document for persistent,
endogenous predictors such as the dividend-price ratio.

Use `cores` to parallelise (the VAR recursion runs in C++, so a replication
costs a few milliseconds; $B = 9999$ takes seconds on a workstation). With
`cores > 1` reproducibility for a given `seed` relies on the L'Ecuyer streams
of the parallel package, so results differ from the serial run.

## Replication of DGRT (2023), Table 4

Table 4, Panel A, tests each Welch–Goyal predictor for the monthly equity
premium, January 1927–December 2020, with asymptotic Eicker–White p-values
and RWB p-values ($B = 9999$). The replication (in the `research/replication`
folder of the source repository, with the data) reproduces the two-sided
p-values for 12 of 14 predictors within bootstrap noise once the *ex-dividend*
S&P return is used as the response; the table lists a selection.

| predictor | EW paper | EW `ivx` | RWB paper | RWB `ivx_boot` |
|---|---|---|---|---|
| dp   | 0.510 | 0.536 | 0.612 | 0.641 |
| ep   | 0.172 | 0.167 | 0.272 | 0.269 |
| bm   | 0.594 | 0.529 | 0.453 | 0.425 |
| tbl  | 0.073 | 0.068 | 0.087 | 0.091 |
| lty  | 0.054 | 0.061 | 0.070 | 0.079 |
| ntis | 0.234 | 0.208 | 0.161 | 0.147 |
| dfy  | 0.996 | 0.985 | 0.994 | 0.979 |
| dfr  | 0.394 | 0.398 | 0.374 | 0.423 |

Two caveats from that exercise. First, the paper's text describes the
response as the return *including* dividends, but the published values for the
three strongly endogenous predictors (dp, ep, bm) are only reproduced with the
ex-dividend series; dy behaves the other way round. Second, the Goyal–Welch
data are revised over time (inflation and net issuance in particular), so
exact agreement is not expected. An independent re-implementation of the
Eicker–White statistic in plain R reproduces the package's C++ values exactly,
so the statistics themselves are verified.

## Caveats

* Bootstrap requires a plain `ivx` fit with `horizon = 1` for the subsample
  tests, but `ivx_boot()` itself accepts any horizon: the null is imposed on
  the one-step response and the long-horizon statistic is recomputed.
* Weighted fits and the `ivx_ar`, `ivx_ra`, `ivx_qr` classes are not supported.
* RWB estimates the predictor's autoregression; under a near-unit root the
  local-to-unity parameter is not consistently estimable, which DGRT show does
  not affect first-order validity but explains why the two schemes differ in
  finite samples.
* On Windows, `cores > 1` starts PSOCK workers that need the package installed.

## References

* Demetrescu, M., Georgiev, I., Rodrigues, P. M. M., & Taylor, A. M. R. (2023).
  Extensions to IVX methods of inference for return predictability. *Journal
  of Econometrics*, 237(2), 105271.
* Hosseinkouchack, M., & Demetrescu, M. (2021). Finite-sample size control of
  IVX-based tests in predictive regressions. *Econometric Theory*, 37(4),
  769–793.
* Welch, I., & Goyal, A. (2008). A comprehensive look at the empirical
  performance of equity premium prediction. *Review of Financial Studies*,
  21(4), 1455–1508.
