---
title: "Rolling IVX tests for bubble detection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rolling IVX tests for bubble detection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4)
```

Pavlidis, Paya & Peel (2017) detect periodically collapsing bubbles by running
the Fama (1984) regression inside a rolling window and drawing inference with
IVX. Everything the procedure needs is already in `ivx()`, so this vignette
shows how to build it in a few lines rather than through a dedicated function.

The procedure is:

1. Fix a window of `w` observations. The paper uses
   `w = max(25, r0 * T)`, where `r0 * T` is the minimum window rule of
   Phillips, Shi & Yu (2015): `r0 = 0.01 + 1.8 / sqrt(T)`.
2. For every window ending at `t = w, ..., T`, fit the predictive regression
   with `ivx()` and keep the IVX t-statistic of the slope. Under the null the
   statistic is standard normal.
3. Compare the sequence against a constant critical value. For the overall
   test of "any bubble in the sample" the paper applies a Bonferroni correction
   for the `T - w + 1` hypotheses, `qnorm(1 - alpha / (T - w + 1))`. For dating
   the episodes it compares each statistic with the plain `qnorm(1 - alpha)`.

```{r}
library(ivx)

ivx_roll <- function(formula, data, window, horizon = 1, alpha = 0.05, ...) {
  n <- nrow(data)
  ends <- window:n
  tstat <- vapply(ends, function(e) {
    fit <- ivx(formula, data[(e - window + 1):e, ], horizon = horizon, ...)
    fit$tstat[1]
  }, numeric(1))
  data.frame(
    end = ends,
    tstat = tstat,
    cv = qnorm(1 - alpha),
    cv_bonf = qnorm(1 - alpha / length(ends))
  )
}
```

`fit$tstat` is `coef / se`, i.e. the IVX t-ratio of the first regressor; the
one-sided alternative in the paper is a slope above its efficient-market value.
With the KMS monthly data:

```{r}
n <- nrow(kms)
w <- max(25, floor((0.01 + 1.8 / sqrt(n)) * n))
r <- ivx_roll(Ret ~ LTY, kms, window = w)

max(r$tstat)
r$cv_bonf[1]
```

The overall test rejects if the maximum statistic exceeds the Bonferroni
critical value. Episodes are dated by the periods where the statistic sits above
the standard normal critical value:

```{r}
plot(r$end, r$tstat, type = "l", xlab = "window end", ylab = "IVX t-statistic")
abline(h = r$cv[1], col = "red", lty = 2)
abline(h = r$cv_bonf[1], col = "red")
```

```{r}
range(r$end[r$tstat > r$cv])
```

For a long-horizon Fama regression pass `horizon = n` and the overlapping
regressand; `ivx()` then uses the Phillips & Lee (2013) long-horizon estimator.
Bootstrap critical values for each window can be obtained by replacing the
`qnorm()` constants with `ivx_boot()` quantiles, at the cost of `B` extra fits
per window.

## References

Pavlidis, E. G., Paya, I., & Peel, D. A. (2017). Testing for speculative
bubbles using spot and forward prices. *International Economic Review*, 58(4),
1191-1226.

Phillips, P. C. B., Shi, S., & Yu, J. (2015). Testing for multiple bubbles:
Historical episodes of exuberance and collapse in the S&P 500. *International
Economic Review*, 56(4), 1043-1078.
