---
title: "Choosing a test: a decision workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choosing a test: a decision workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
has_qr <- requireNamespace("quantreg", quietly = TRUE)
```

```{r setup, message = FALSE}
library(ivx)
```

The package holds a dozen estimators and tests. Each vignette documents one of
them; this one is the map. It runs a predictive regression from the plain
`ivx()` fit through the diagnostics that decide which extension applies, and
follows every branch with the `kms` (monthly S&P 500 excess returns, Kostakis
et al. 2015) and `ylpc` (quarterly US housing returns, Yang et al. 2020) data.

## The map

```
ivx(y ~ x, data)                       KMS Wald test, the baseline
 │
 ├─ residuals serially correlated?  ac_test()          → ivx_ar()
 ├─ residuals heteroskedastic?      ac_test(res^2)     → ivx(robust = TRUE)
 ├─ |delta| near 1 and x near a unit root, short sample, one-sided test?
 │                                   delta(), hlt_test  → ivx_boot(), ivx_ra()
 ├─ weak signal, power a concern?                      → ivx(lag_y = TRUE)
 ├─ K-period cumulative return?                        → ivx(horizon = K), ivx_ra(horizon = K)
 ├─ several responses, cross-equation restrictions?    → ivx_sys()
 ├─ tail or median predictability?                     → ivx_qr(), ivx_qr_boot()
 ├─ predictability confined to episodes?               → ivx_episodic(), rolling ivx()
 └─ single predictor, want a non-IVX cross-check?      → hlt_test(), cy_test(), arm(), el_test(), ivx_iv()
```

The branches are not exclusive: the worked example below ends up combining
three of them.

## Step 1: baseline fit and three diagnostics

```{r}
m <- ivx(Ret ~ DP + TBL, data = kms)
summary(m)
```

The IVX Wald test is valid for any degree of persistence in $x_t$, so the
first question is not "is $x_t$ persistent?" but "is anything else wrong with
the errors, and is the sample short enough for finite-sample distortions to
matter?". Three checks answer it.

**Persistence and endogeneity.** The lag-1 autocorrelation of each predictor
and the correlation $\delta$ between the return innovation and the predictor
innovation:

```{r}
sapply(kms[c("DP", "TBL")], function(x) cor(head(x, -1), tail(x, -1)))
delta(m)
```

Both predictors are near unit roots. `DP` is also strongly endogenous,
$\delta \approx `r round(delta(m)[1], 2)`$, the combination under which the
asymptotic test over-rejects in samples of a few hundred observations
(Hosseinkouchack & Demetrescu 2021; Demetrescu et al. 2023). `TBL` is
persistent but nearly exogenous, so its asymptotic p-value can be taken at
face value. `hlt_test()` classifies persistence formally through the ADF
normalised-bias statistic (see its `adf` and `test` components) if a rule is
preferred to eyeballing.

**Serial correlation in the errors.** `ac_test()` on the fit reports four
tests at lags 1 to `lag_max`. The `Wald` column is the regression-based test
of Yang et al. (2020), the one `ivx_ar()` reports as `Wald_AR`; the
Box tests are portmanteau tests on the residual autocorrelations.

```{r}
ac_test(m, 4)
```

The Box and Breusch–Godfrey statistics are significant, the Wald test is
not. The first-order autocorrelation of the residuals is
`r round(acf(residuals(m), plot = FALSE)$acf[2], 3)`: the mild
low-order dependence typical of monthly returns, which is not what
`ivx_ar()` is for. Compare the housing data:

```{r}
mh <- ivx(hpi ~ log(res) + cpi, data = ylpc)
ac_test(mh, 4)
```

Every statistic is an order of magnitude larger and the Wald test rejects at
lag 1. That is the `ivx_ar()` case (Step 2).

**Heteroskedasticity.** The same portmanteau tests on the squared residuals
are the McLeod–Li test for ARCH effects:

```{r}
ac_test(residuals(m)^2, 4)
```

Strong volatility clustering, as expected for returns. The Eicker–White
standard errors of Step 3 handle it.

## Step 2: serially correlated errors → `ivx_ar()`

Serial correlation in $u_t$ breaks the $\chi^2$ limit of the IVX Wald
statistic. `ivx_ar()` fits an AR($q$) to the residuals, quasi-differences the
regression over a grid of AR coefficients and reports the IVX statistics of
the fit with the smallest residual variance (`vignette("ivx-ar")`).

```{r}
ma <- ivx_ar(hpi ~ log(res) + cpi, data = ylpc)
summary(ma)
```

`Wald_AR` is the test of $\phi_1 = \dots = \phi_q = 0$; when it does not
reject, the profiling adds noise and plain `ivx()` is the better choice.
`ivx_ar()` passes `robust`, `beta`, `cz` and `bandwidth` through to every grid
refit, but `ivx_boot()` does not accept its output: the bootstrap is defined
for the untransformed regression.

## Step 3: heteroskedasticity → `robust = TRUE`

Demetrescu, Georgiev, Rodrigues & Taylor (2023) show the IVX statistics keep
their limits under conditional and unconditional heteroskedasticity if the
homoskedastic $\hat\sigma^2_u \sum z z'$ is replaced by the Eicker–White
$\sum z_{t-1} z_{t-1}' \hat u_t^2$. There is no cost when the errors are
homoskedastic, so for return data this is the sensible default.

```{r}
mr <- ivx(Ret ~ DP + TBL, data = kms, robust = TRUE)
coef(summary(mr))
```

`robust = TRUE` is defined for `horizon = 1`; `ivx_ra()` uses
heteroskedasticity-robust standard errors at every horizon.

## Step 4: finite-sample size → `ivx_boot()` and `ivx_ra()`

For a strongly endogenous, near-unit-root predictor such as `DP`, the
asymptotic test is oversized in finite samples, more so one-sided. Two
remedies, in order of preference:

1. **Bootstrap the statistic.** The residual wild bootstrap of DGRT (2023)
   refits the predictor's autoregression, resamples $(u_t, v_t)$ with a
   common multiplier so $\delta$ is preserved, and recomputes the IVX
   statistics (`vignette("robust-inference")`).

```{r}
mb <- ivx_boot(m, B = 499, type = "rwb", seed = 1)
mb
```

2. **Bias-reduce the estimator.** `ivx_ra()` (Demetrescu & Rodrigues 2022)
   augments the regression with the residuals of the predictors'
   autoregression, which removes the endogeneity-driven bias before
   instrumenting. Its standard errors are robust by construction.

```{r}
mra <- ivx_ra(Ret ~ DP + TBL, data = kms)
coef(summary(mra))
```

The four p-values for the two-sided test of `DP` line up as the theory
predicts, with the asymptotic homoskedastic one the most optimistic:

```{r}
p_dp <- c(
  asymptotic = coef(summary(m))["DP", "Pr(> chi)"],
  eicker_white = coef(summary(mr))["DP", "Pr(> chi)"],
  rwb_bootstrap = mb$p.value$Wald_Ind[["DP"]],
  residual_augmented = coef(summary(mra))["DP", "Pr(> chi)"]
)
round(p_dp, 3)
```

Use `cores > 1` for the bootstrap with $B$ in the thousands; use
`ivx_boot(type = "frwb")` if the predictor's autoregression is unstable
(very short samples, structural breaks in the predictors).

## Step 5: power → `lag_y = TRUE`

The IVX instrument is less persistent than $x_t$, so the test gives up local
power against OLS exactly when $x_t$ is near-integrated and endogenous.
Demetrescu (2014) recovers part of it by adding $y_{t-1}$ to the regression,
instrumented by itself. This is a modelling choice rather than a diagnostic
outcome: it costs nothing asymptotically when the extra term is irrelevant
and can raise power substantially when the signal is weak.

```{r}
coef(summary(ivx(Ret ~ DP + TBL, data = kms, lag_y = TRUE)))
```

The joint Wald statistic still tests only the predictors.

## Step 6: what is the null?

The steps above fix the error structure. The remaining branches depend on
what is being tested.

**Long horizons.** For a $K$-period cumulative return, `horizon = K` gives
the Kostakis et al. (2023) statistic, which handles the overlap of the sums
through the instrument. `ivx_ra(horizon = K)` is the transformed-regression
alternative of Demetrescu, Rodrigues & Taylor (2023), which avoids HAC
estimation altogether. Report both when the horizon is long relative to the
sample: they use different information and the overlap ($K - 1$
observations per statistic) is where long-horizon tests misbehave.

```{r}
c(kms = ivx(Ret ~ DP + TBL, data = kms, horizon = 12)$Wald_Joint,
  drt = ivx_ra(Ret ~ DP + TBL, data = kms, horizon = 12)$Wald_Joint)
```

**Several responses.** `ivx_sys()` (Magdalinos 2022) estimates the system
jointly and returns the full covariance of $\mathrm{vec}(A)$, so restrictions
across equations, e.g. "does `DP` predict the return and the payout ratio
equally?", are Wald tests on `vcov()` (`vignette("ivx-sys")`).

**Quantiles.** If the question is about the tails ("does `DP` predict large
losses?") rather than the mean, `ivx_qr()` (Lee 2016) runs the IVX test in a
quantile regression; `ivx_qr_boot()` gives block-bootstrap intervals that
are robust to the conditional heteroskedasticity found in Step 1
(`vignette("ivx-qr")`).

```{r, eval = has_qr}
q <- ivx_qr(Ret ~ DP + TBL, data = kms, tau = c(0.1, 0.5, 0.9))
sapply(q, function(fit) c(Wald_Joint = fit$Wald_Joint, p = 1 - pchisq(fit$Wald_Joint, 2)))
```

**Episodes.** A full-sample rejection can be driven by a short pocket of
predictability, and a full-sample non-rejection can hide one.
`ivx_episodic()` computes the IVX statistic over rolling, forward or backward
windows with the full-sample instrument and bootstraps the supremum of the
sequence (`vignette("ivx-episodic")`); a plain rolling `ivx()` with
Bonferroni critical values is the bubble-detection variant of
`vignette("rolling-ivx")`.

```{r}
ivx_episodic(ivx(Ret ~ DP, data = kms), scheme = "rolling", window = 0.2, B = 199, seed = 1)
```

## Step 7: cross-check with a non-IVX test

With a single predictor, the benchmarks in `vignette("benchmarks")` answer
the same question from a different direction and are cheap to run. The
hybrid test of Harvey, Leybourne & Taylor (2021) is the most useful
companion: it classifies the predictor's persistence, picks the test
accordingly and reports the intermediate quantities.

```{r}
hlt_test(Ret ~ DP, data = kms)
cy_test(Ret ~ DP, data = kms)
```

`cy_test()` (Campbell & Yogo 2006) gives a Bonferroni confidence interval for
$\beta$ that is exact for a near-unit-root predictor but conservative
otherwise; `arm()` and `el_test()` are the reduced-bias OLS and empirical
likelihood alternatives; `ivx_iv()` (Breitung & Demetrescu 2015) replaces the
IVX filter by sine or fractional-difference instruments, a check that the
result does not hinge on the choice of $\rho_n$.

## The worked example, in one table

For `Ret ~ DP + TBL` the diagnostics said: near-unit-root predictors, `DP`
strongly endogenous, mild residual autocorrelation, strong ARCH. The
decisions that follow are Eicker–White standard errors, a bootstrap p-value
for the endogenous predictor, and `ivx_ra()` as the bias-reduced
cross-check; no `ivx_ar()`. The p-values collected above, plus the joint
test:

```{r}
rbind(
  DP = round(p_dp, 3),
  joint = round(c(
    asymptotic = 1 - pchisq(m$Wald_Joint, 2),
    eicker_white = 1 - pchisq(mr$Wald_Joint, 2),
    rwb_bootstrap = mb$p.value$Wald_Joint,
    residual_augmented = 1 - pchisq(mra$Wald_Joint, 2)
  ), 3)
)
```

Every robust version moves the p-value up, and the more the method does
about endogeneity the further it moves: the pattern DGRT (2023) document for
the dividend–price ratio. Here nothing rejects at any conventional level, but
with a borderline asymptotic p-value the ordering is what decides the
conclusion.

For `hpi ~ log(res) + cpi` the diagnostics said: strong residual
autocorrelation. The decision is `ivx_ar()`, whose `Wald_AR` confirms the
AR terms and whose IVX statistics replace the plain ones.

## Summary

| Symptom | Diagnostic | Use |
|---|---|---|
| Serially correlated errors | `ac_test(fit)`, `Wald` column | `ivx_ar()` |
| Conditional heteroskedasticity | `ac_test(residuals(fit)^2)` | `ivx(robust = TRUE)` |
| $\lvert\deltavert$ near 1, $x$ near unit root, short sample, one-sided test | `delta()`, `hlt_test()$adf` | `ivx_boot(type = "rwb")`, `ivx_ra()` |
| Weak signal | — | `ivx(lag_y = TRUE)` |
| Multi-period return | — | `ivx(horizon = K)`, `ivx_ra(horizon = K)` |
| Several responses | — | `ivx_sys()` |
| Tail predictability | — | `ivx_qr()`, `ivx_qr_boot()` |
| Pockets of predictability | plot of rolling `ivx()` t-ratios | `ivx_episodic()` |
| Single predictor, second opinion | — | `hlt_test()`, `cy_test()`, `arm()`, `el_test()`, `ivx_iv()` |

## References

* Breitung, J., & Demetrescu, M. (2015). Instrumental variable and variable
  addition based inference in predictive regressions. *Journal of
  Econometrics*, 187(1), 358–375.
* Campbell, J. Y., & Yogo, M. (2006). Efficient tests of stock return
  predictability. *Journal of Financial Economics*, 81(1), 27–60.
* Demetrescu, M. (2014). Enhancing the local power of IVX-based tests in
  predictive regressions. *Economics Letters*, 124(2), 269–273.
* Demetrescu, M., Georgiev, I., Rodrigues, P. M. M., & Taylor, A. M. R. (2023).
  Extensions to IVX methods of inference for return predictability. *Journal
  of Econometrics*, 237(2), 105271.
* Demetrescu, M., & Rodrigues, P. M. M. (2022). Residual-augmented IVX
  predictive regression. *Journal of Econometrics*, 227(2), 429–460.
* Demetrescu, M., Rodrigues, P. M. M., & Taylor, A. M. R. (2023). Transformed
  regression-based long-horizon predictability tests. *Journal of
  Econometrics*, 237(2), 105316.
* Harvey, D. I., Leybourne, S. J., & Taylor, A. M. R. (2021). Simple tests for
  stock return predictability with good size and power properties. *Journal
  of Econometrics*, 224(1), 198–214.
* Hosseinkouchack, M., & Demetrescu, M. (2021). Finite-sample size control of
  IVX-based tests in predictive regressions. *Econometric Theory*, 37(4),
  769–793.
* Kostakis, A., Magdalinos, T., & Stamatogiannis, M. P. (2015). Robust
  econometric inference for stock return predictability. *Review of Financial
  Studies*, 28(5), 1506–1553.
* Kostakis, A., Magdalinos, T., & Stamatogiannis, M. P. (2023). Taking stock
  of long-horizon predictability tests: Are factor returns predictable?
  *Journal of Econometrics*, 237(2), 105380.
* Lee, J. H. (2016). Predictive quantile regression with persistent
  covariates: IVX-QR approach. *Journal of Econometrics*, 192(1), 105–118.
* Magdalinos, T. (2022). Least squares and IVX limit theory in systems of
  predictive regressions with GARCH innovations. *Econometric Theory*, 38(5),
  875–912.
* Yang, B., Long, W., Peng, L., & Cai, Z. (2020). Testing the predictability
  of US housing price index returns based on an IVX-AR model. *Journal of the
  American Statistical Association*, 115(532), 1598–1619.
